The Chemistry Development Kit (CDK): atom typing, rendering, molecular formulas, and substructure searching

### 2016-10-09 article.tex
More details about the fingerprints

\subsubsection*{Molecular Fingerprints}
Molecular fingerprints are another area that seen significant
development. Traditionally fingerprints were represented using the
\texttt{BitSet} class from the Java library. While using this class
allowed the use of pre-existing methods to manipulate bit strings, it
keeps a vector of bits in memory. The solution was excellent for
hashed, relatively small fingerprints, \textit{e.g.}, 1024 bits,
\textit{i.e.}, 10 bits indexing space. However implementing a
fingerprint designed to avoid collisions with, \textit{e.g.} a 32 bit
indexing space using this approach would be highly
memory-inefficient. To allow for multiple fingerprint representation
implementations a bit fingerprint interface was introduced\todo{is
  there any more information about this?}. Also,
although fingerprints traditionally are bit vectors a count
fingerprint was also introduced making fingerprints based on integer
vectors supported in CDK as well. The fingerprints currently existing
in CDK are listed in Table~3\ref{tab:fingerprints}.

>>>>

\subsubsection*{Molecular Fingerprints}
Molecular fingerprints have also seen significant development in this CDK version.
Previously, fingerprints were represented using the \texttt{BitSet} class from
the Java library. While using this class
allowed the use of pre-existing methods to manipulate bit strings, it
keeps a vector of bits in memory. The solution was excellent for
hashed, relatively small fingerprints, \textit{e.g.}, 1024 bits,
\textit{i.e.} with a 10 bits indexing space. However, implementing a
fingerprint designed to avoid collisions with, \textit{e.g.} a 32 bit
indexing space using this approach would be highly
memory-inefficient. To allow for multiple fingerprint representation
implementations a bit fingerprint interface was introduced:
\texttt{IBitFingerprint}.

\vspace{0.2cm}
\begin{verbatim}
IFingerprinter fingerprinter = new Fingerprinter();
IBitFingerprint fingerprint = fingerprinter.getBitFingerprint(mol);
\end{verbatim}
\vspace{0.2cm}

Also, although fingerprints traditionally are bit vectors a count
fingerprint was also introduced making fingerprints based on integer
vectors supported in CDK as well. This is useful for fingerprints that are based
on substructures, such as the \texttt{SignatureFingerprinter} which uses the
aforementioned atomic signatures. The counts in the fingerprint then represent
how often this substructure is found in the molecule it represents. For
example, for signature depth 1:

\vspace{0.2cm}
\begin{verbatim}
SignatureFingerprinter fingerprinter = new SignatureFingerprinter(2);
ICountFingerprint bitFP = fingerprinter.getCountFingerprint(mol);
\end{verbatim}
\vspace{0.2cm}

The fingerprints currently provided by the CDK are listed in Table~3\ref{tab:fingerprints}.
